---
title: "Lab 4 Submission"
author: "Sean Kim"
format:
  html:
    embed-resources: true
editor: 
  markdown: 
    wrap: 72
---

```{r}

if (!file.exists("met_all.gz")) 
  download.file( 
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", 
    destfile = "met_all.gz", 
    method = "libcurl", 
    timeout = 60 )
library(data.table)
library(leaflet)
library(tidyverse)
library(lubridate)

met <- data.table::fread("met_all.gz") 
```

2.  Preparing the data:

    ```{r}
    summary(met)
    met <- met[met$temp > -17][elev == 9999.0, elev:= NA]
    met[, date := as.Date(paste(year, month, day, sep = "-"))]

    met.first <- met[day(date) %in% 1:7, ]

    met_avg <- met[,.(
      temp = mean(temp,na.rm=TRUE), 
      rh = mean(rh,na.rm = TRUE), 
      wind.sp = mean(wind.sp,na.rm=TRUE), 
      vis.dist = mean(vis.dist,na.rm=TRUE),
      dew.point = mean(dew.point,na.rm=TRUE), 
      lat = mean(lat), 
      lon = mean(lon), 
      elev = mean(elev,na.rm=TRUE)
    ), by = c("USAFID", "day")]

    met_avg[, region := ifelse(lon <= -98.00 & lat >= 37.71, "NW", 
                                ifelse(lon <= -98.00 & lat <39.71, "SW",
                                       ifelse(lon > -98.00 & lat >= 39.71, "NE", "SE")))]

    met_avg[, elev_cat := ifelse(elev >252, "high", "low")]
    ```

3.  Use geom_violin to examine wind speed and dew point by region

    ```{r}
    met_avg[!is.na(elev_cat)] %>% 
      ggplot()+
      geom_violin(mapping = aes(x = wind.sp, y = dew.point))+
      facet_wrap(~region, nrow = 2)
    ```

4.  Use geom_jitter with stat_smooth to examine the association between
    dew point and wind speed by region.

    ```{r}
    ggplot(met_avg, aes(x = dew.point, y = wind.sp))+
      facet_wrap(~region, nrow = 2)+
      geom_jitter(width = 0.2, height = 0.2) + 
      stat_smooth(method = "lm", se = FALSE) + 
      labs(title = "Association btw Dew Point and Wind Speed by Region", x = "Dew Point", y = "Wind Speed", color = "region") + 
      theme_bw()
    ```

increasing dew point was associated with slight increasing wind speed.

5.  Use geom_bar to create barplots of the weather stations by elevation
    category colored by region

    ```{r}
    met_avg[!is.na(elev_cat)] %>% 
      ggplot()+
      geom_bar(aes(elev_cat, fill = region), position = "dodge") + 
      labs(title = "Weather stations by Elevation Category, colored by region", 
           x = "elevation category", 
           y = "count")
    ```

6.  Use stat_summary to examine mean dew point and wind speed by region
    with standard deviation error bars

    ```{r}
    ggplot(met_avg, aes(x = region, y = dew.point)) +
      stat_summary(fun.data = "mean_sdl", mult = 1, geom = "bar") +
      stat_summary(fun.data = "mean_sdl", mult = 1, geom = "errorbar", width = 0.2) +
      labs(title = "Mean Dew Point by Region with Standard Deviation Error Bars",
           x = "Region",
           y = "Mean Dew Point") +
      theme_minimal()

    ggplot(met_avg, aes(x = region, y = wind.sp))+
      stat_summary(fun.data = "mean_sdl", mult = 1, geom = "bar") + 
      stat_summary(fun.data = "mean_sdl", mult = 1, geom = "errorbar", width = 0.2) + 
      labs( title = "Mean Wind Speed by Region with Standard Deviation Error Bars", 
            x = "Region",
            y = "Mean Wind Speed")+
      theme_minimal()
    ```

7.  make a map showing the spatial trend in relative humidity in the us

    ```{r}
    library(leaflet)

    met<- met[!is.na(rh)]
    colors <- colorNumeric(palette = "viridis", domain = met$rh)
    map <- leaflet(data = met) %>%
      addTiles() %>%
      addCircleMarkers(~lon, ~lat, radius = 4, color = ~colors(rh),
                       popup = ~paste("USAFID: ", USAFID, "<br>Relative Humidity: ", rh, "%")) %>%
      addLegend(pal = colors, values = ~rh, title = "Relative Humidity (%)") %>%
      addMarkers(data = met[order(-rh)][1:10, ], ~lon, ~lat, label = ~USAFID)

    ```

8.  use a ggplot extension - ggthemes

    ```{r}
    library(ggthemes)
    ggplot(met_avg, aes(x = region, y = dew.point)) + 
      geom_violin() + 
      facet_wrap(~region) + 
      labs(title = "dew point by region violin plot tufte theme", 
          x = "region", 
          y = "dew point")+ 
      theme_tufte()
    ```
